---
title: "Gapminder: 200 Countries, 200 Years"
subtitle: "Capstone project: BIDA 302"
author: "Matheus Buarque"


output: html_document
---


$~$
$~$
$~$


### These are the packeges I used on this project

Package load:

```{r setup, message=FALSE}

library(dplyr)
library(tidyverse)
library(readxl)
library(ggplot2)
options(scipen=999)
install.packages('gganimate')


```

$~$

### Here's where I start importing the data using read() function, pivot_longer(), and mutate() to avoid problems when plotting.

```{r warning=FALSE, message=FALSE}
gdp <- read_csv("data/income_per_person_gdppercapita_ppp_inflation_adjusted.csv") %>% 
  pivot_longer(-country, names_to = "year", values_to = "income_person") %>% 
  mutate(year = as.integer(year)) #here I read, pivot, and clean my data at once assignt it to gdp (income)

gdp


life_expectancy <- read_csv("data/life_expectancy_years.csv") %>% 
  pivot_longer(-country, names_to = "year", values_to = "life_exp") %>% 
  mutate(year = as.integer(year))

life_expectancy

population <- read_csv("data/population_total.csv") %>% 
  pivot_longer(-country, names_to = "year", values_to = "population") %>% 
  mutate(year = as.integer(year)) %>% 
  mutate(population = as.double(population))

population


geo_location <- read_excel("data/Data Geographies - v1 - by Gapminder.xlsx", sheet = "list-of-countries-etc") %>% 
  select("name", "four_regions") %>% 
  dplyr::rename(country = name)

geo_location


```

$~$
$~$
$~$


### Here I join the tables using left_join() function.

```{r }

table1 <- geo_location %>% 
  left_join(gdp,by = "country") %>% 
  left_join(life_expectancy, by = c("country","year")) %>% 
  left_join(population, by = c("country","year"))

```

$~$
$~$
$~$

### With ggplot() I start creating and custumizing my plot

```{r warning=FALSE}

table1810 <- table1 %>% 
  filter(year == "1810",
         na.rm = TRUE) %>% #to avoid NA values
ggplot(aes(x = income_person, 
             y = life_exp, 
             colour = four_regions,
             size = population)) +
    geom_point() +
    labs(title = "Life expectancy",
       subtitle = "1810",
       caption = "Source: Gapminder",
       y = "Life expectancy",
       x = "Income",
       color = "Continent",
       size = "Population") + #adding labels to my plot
    scale_y_continuous(
      limits = c(25, 75),
      breaks = c(25, 50, 75),
      minor_breaks = NULL) + #minor_braks remove the unnecessary limits on my scale
    scale_x_continuous(
      limits = c(400, 40000),
      breaks = c(400, 4000, 40000),
      minor_breaks = NULL,
      trans = "log10") #the argument trans = "log10" multiply my y axis by 10 (400, 4000, 40000), giving them the
                       #space ("block")
  
table1810

table1996 <- table1 %>% 
  filter(year == "1996",
         na.rm = TRUE) %>% 
ggplot(aes(x = income_person, 
             y = life_exp, 
             colour = four_regions,
             size = population)) +
    geom_point()+
    labs(title = "Life expectancy",
       subtitle = "1996",
       caption = "Source: Gapminder",
       y = "Life expectancy",
       x = "Income",
       color = "Continent",
       size = "Population") +
    scale_y_continuous(
      limits = c(25, 75),
      breaks = c(25, 50, 75),
      minor_breaks = NULL) +
    scale_x_continuous(
      limits = c(400, 40000),
      breaks = c(400, 4000, 40000),
      minor_breaks = NULL,
      trans = "log10")

table1996

table2009 <- table1 %>% 
  filter(year == "2009",
         na.rm = TRUE) %>%  
ggplot(aes(x = income_person, 
             y = life_exp, 
             colour = four_regions,
             size = population)) +
    geom_point() +
    labs(title = "Life expectancy",
       subtitle = "2009",
       caption = "Source: Gapminder",
       y = "Life expectancy",
       x = "Income",
       color = "Continent",
       size = "Population") +
    scale_y_continuous(
      limits = c(25, 75),
      breaks = c(25, 50, 75),
      minor_breaks = NULL) +
    scale_x_continuous(
      limits = c(400, 40000),
      breaks = c(400, 4000, 40000),
      minor_breaks = NULL,
      trans = "log10")
  

table2009


```

$~$
$~$

### And last, I use gganimate and scales to create a gif plot, simpler than Hans Rosling but with the same idea of showing the progress of life-expectancy over the years.

```{r warning=FALSE, message=FALSE}
library(gganimate)
library(scales)

g <- ggplot(table1, aes(x = income_person, 
             y = life_exp, 
             colour = four_regions,
             size = population,
             label = as.character(round(year, 2)))) + #here I just copied the labels, but used
                                                     #as.character(round(year, 2)) to hide decimals in my title
  geom_point(show.legend = TRUE, alpha = 0.17) + #here I enable the legends
  scale_color_viridis_d() +
  scale_size(range = c(2, 15)) +
  labs(caption = "Source: Gapminder",
       y = "Life expectancy",
       x = "Income",
       color = "Continent",
       size = "Population") +
  scale_y_continuous(
      limits = c(25, 75),
      breaks = c(25, 50, 75),
      minor_breaks = NULL) +
  scale_x_continuous(
      limits = c(400, 40000),
      breaks = c(400, 4000, 40000),
      minor_breaks = NULL,
      trans = "log10")

g + transition_time(year) +
  labs(title = "Year: {frame_time}") #transition_time(year) create a gif animation, layer over layer according to year



```